Constructive Preference Elicitation by Setwise Max-margin Learning

### 2016-01-13 paper/ijcai16.tex
paper: add some meat to the experiments + random updates.

In this section we provide extensive experimental evaluation of the {\sc
SetMargin} algorithm. We implemented the {\sc SetMargin} algorithm using
Python, and leverage Gurobi 6.5.0 for solving the MILP problem. Both the 
{\sc SetMargin} source code and the full experimental setup are available at
\stefano{add URL}.

@@

We adopted the {\em indifference-augmented} Bradley-Torr user response model
introduced in~\cite{guo2010real}. When queried about a pair of configurations,
the user replies with the true utility of the two items. According to the
classical Bradley-Torr model, which does not attempt to model {\em indifference},
the proabability that a user $\vw$ will prefer configuration $\vx^i$ over
the alternative $\vx^j$, barring indifference, amounts to:
%
$$ P(\vx^i > \vx^j \;|\; \vw, \vx^i \not\sim \vx^j) = (1 + \exp(-\alpha \langle\vw,\vx^i - \vx^j\rangle))^{-1} $$
%
In other words, the probability grows with the difference between the utilities
of the two items. Support for indifference has been introduced
in~\cite{guo2010real} by postulating that the probability of the user being
indifferent about the two configurations $\vx^i, \vx^j$ depends on how close
the two utilities are, as in:
%
$$ P(\vx^i \sim \vx^j \;|\; \vw) = \exp(\beta |\langle\vw,\vx^i - \vx^j\rangle|) $$
%
The parameters $\alpha$ and $\beta$ were set as in~\cite{guo2010real}. Please
see the latter for more details
\stefano{The above does not satisfy me at all.}

In all experiments we use an internal cross-validation procedure to update the
hyperparameters $\alpha$, $\beta$, and $\gamma$ after every 5 iterations. The
hyperparameters are chosen as to maximize the ranking loss over the user
answers collected so far. $\alpha$ is taken in $\{100, 10, 1\}$,
while $\beta$ and $\gamma$ are taken in $\{100, 10, 1, 0.1, 0\}$. Note that
$\beta=0$ implies that the weight vector is not regularized, nor sparse, while
$\gamma=0$ implies that the absolute quality of the configurations is ignored.
\stefano{is this desirable?}

\paragraph{Synthetic Dataset.} Following the experimental protocol
in~\cite{guo2010real} and \cite{viappiani2010optimal}, in the first experiment
we evaluate the behavior of the proposed method in a challenging artificial
setting. We developed seven synthetic datasets: each dataset involves
$n=2,\ldots,7$ attributes, where each attribute takes one of $n$ possible
values. More explicitly, for $n=2$ the synthetic dataset is
$\calX_\text{feasible} = [2] \times [2]$, for $n=3$ to $\calX_\text{feasible} =
[3] \times [3] \times [3]$, and so on. The cardinality of
$\calX_\text{feasible}$ is $n^n$, and grows (super-)exponentially with $n$. For
$n=3$, the dataset is comparable to the synthetic one used
in~\cite{guo2010real} and~\cite{viappiani2010optimal} (which have three domains
of size of 2, 2 and 5, respectively, for a total of 20 feasible
configurations), while for larger $n$ it grows much larger than those typically
used in the Bayesian preference elicitation literature, and as such represents
a good test bed for comparing the scalability of the various methods.
The feasible configuration space was encoded in {\sc SetMargin} through
appropriate MILP constraints, while the other methods required all datasets to
be explicitly grounded.

Users were simulated by drawing random utility vectors from four different
distributions. The first two mimic those used in~\cite{guo2010real}: (1) a
uniform distribution over $[1, 100]$ for each individual weight, and (2) a
normal distribution with mean $25$ and covariance $\frac{25}{3}$\footnote{This
procedure differs from the one reported in~\cite{guo2010real}; we opted for
following the procedure implemented in their code. \stefano{more diplomacy needed}}.
We further produced two novel {\em sparse} sets of weight vectors by deleting
$80\%$ uniformly at random \stefano{Paolo, double check please} the entries
from the vectors sampled in the uniform and normal settings. We sampled $100$
vectors for each setting. All methods were evaluated on the very same weight
vectors.

@@

\paragraph{Constructive Dataset.} Next, we tested {\sc SetMargin} on a truly
constructive setting, which can not be readily handled by non-constructive
methods.

The attributes are Manufacturer, CPU, Monitor size, PC type, Memory, and HD
size. CPU type and CPU frequency as combined in a single variable, e.g. ``AMD
Athlon @1000" stands for an AMD Athlon CPU clocked at 1 GHz. All other domains
are left unchanged.

The configurations constraints:
%
$$ x_i \Leftarrow x_{j_1} \lor \ldots \lor x_{j_n} $$
%
have been translated into MILP constraints of the form:
%
$$ (1 - x_i) + x_{j_1} + \ldots + x_{j_n} \ge 1 $$

Each attribute configuration is associated a corresponding {\em price}. The
price of a configuration is the sum of the prices of the individual components.

\begin{figure}
    {\footnotesize
    \begin{tabular}{cc}
        {\bf Rule} & {\bf MILP constraint} \\ \hline\hline
        & Manufacturer = Compacq $\Rightarrow$ Type = Laptop $\lor$ Type = Desktop \\
        & Manufacturer = Apple $\Rightarrow$ CPU = PowerPC G3 @266 $\lor$ \ldots \\
        & Type = Laptop $\Rightarrow$ Memory = 64 $\lor$ \ldots \\
        & Type = HDSize $\Rightarrow$ HDSize = 20 $\lor$ \ldots \\
        & Type = Desktop $\Rightarrow$ Monitor = 15 $\lor$ \ldots \\
    \end{tabular}
    }
    \caption{\label{tab:pcconstraints} List of MILP constraints used in the
    PC experiment.}
\end{figure}

>>>>

In this section we provide extensive experimental evaluation of the {\sc
SetMargin} algorithm. We implemented the {\sc SetMargin} algorithm using
Python, leveraging Gurobi 6.5.0 as the MILP solver. Both the {\sc SetMargin}
source code and the full experimental setup are available at \stefano{add URL}.

@@

We adopted the {\em indifference-augmented} Bradley-Torr user response model
introduced in~\cite{guo2010real}. When queried about a pair of configurations,
the user replies with the true utility of the two items. According to the
classical Bradley-Torr model, which does not attempt to model {\em indifference},
the proabability that a user $\vw$ will prefer configuration $\vx^i$ over
the alternative $\vx^j$, barring indifference, amounts to:
%
$$ (1 + \exp(-\alpha \langle\vw,\vx^i - \vx^j\rangle))^{-1} $$
%
In other words, the probability grows with the difference between the utilities
of the two items. Support for indifference has been introduced
in~\cite{guo2010real} by postulating that the probability of the user being
indifferent about the two configurations $\vx^i, \vx^j$ depends on how close
the two utilities are, as in:
%
$$ \exp(\beta |\langle\vw,\vx^i - \vx^j\rangle|) $$
%
The parameters $\alpha$ and $\beta$ were set as in~\cite{guo2010real}. Please
see the latter for more details
\stefano{The above does not satisfy me at all.}

In all experiments we use an internal 5-fold cross-validation procedure to update the
hyperparameters $\alpha$, $\beta$, and $\gamma$ after every 5 iterations. The
hyperparameters are chosen as to minimize the ranking loss over the user
answers collected so far. $\alpha$ is taken in $\{20, 15, 10, 5, 1\}$,
while $\beta$ and $\gamma$ are taken in $\{10, 5, 1, 0.5, 0.1\}$.
\stefano{explain that we can't turn off $\gamma$}

\paragraph{Synthetic Dataset.} Following the experimental protocol
in~\cite{guo2010real} and \cite{viappiani2010optimal}, in the first experiment
we evaluate the behavior of the proposed method in a challenging artificial
setting. We developed seven synthetic datasets: each dataset involves
$n=2,\ldots,7$ attributes, where each attribute takes one of $n$ possible
values. More explicitly, for $n=2$ the synthetic dataset is
$\calX_\text{feasible} = [2] \times [2]$, for $n=3$ to $\calX_\text{feasible} =
[3] \times [3] \times [3]$, and so on. The cardinality of
$\calX_\text{feasible}$ is $n^n$, and grows (super) exponentially with $n$. For
$n=3$, the dataset is comparable to the synthetic one used
in~\cite{guo2010real} and~\cite{viappiani2010optimal} (which have three domains
of size of 2, 2 and 5, respectively, for a total of 20 feasible
configurations), while for larger $n$ it grows much larger than those typically
used in the Bayesian preference elicitation literature, and as such represents
a good test bed for comparing the scalability of the various methods.
The feasible configuration space was encoded in {\sc SetMargin} through
appropriate MILP constraints, while the other methods required all datasets to
be explicitly grounded.

Users were simulated by drawing random utility vectors from four different
distributions. The first two mimic those used in~\cite{guo2010real}: (1) a
uniform distribution over $[1, 100]$ for each individual weight, and (2) a
normal distribution with mean $25$ and covariance $\frac{25}{3}$\footnote{This
procedure differs from the one reported in~\cite{guo2010real}; we opted for
following the procedure implemented in their code. \stefano{more diplomacy needed}}.
We further produced two novel {\em sparse} sets of weight vectors by deleting
$80\%$ of the entries of the vectors sampled in the uniform and normal settings
uniformly at random \stefano{Paolo, is this correct?}. We sampled $100$
vectors for each setting. All methods were evaluated on the very same weight
vectors.

@@

\paragraph{Constructive dataset.} Next, we tested {\sc SetMargin} on a truly
constructive setting. We developed a constructive version of the PC dataset
used in~\cite{guo2010real}: instead of explicitly enumerating all possible PC
item, we definined the set of feasible configurations through MILP constraints.

A PC configurations is defined by seven attributes: computer type (laptop,
desktop, or tower), manufacturer, CPU model, monitor size, RAM amount, storage
size, and price. See Table~\ref{tab:pcdataset} for an overivew of the
attributes. The price attribute is defined as a linear combination of the other
attributes: this is a fair modeling choice, as often the price of a PC is
well approximated by sum of the price of its components plus an offset due to
branding. Compared to the PC dataset in~\cite{guo2010real}, we combined the CPU
type and CPU frequency attributes into a single CPU attribute (e.g. ``PowerPC
G4 @733 MHz``), so to avoid multiplicative terms in the price function.
\stefano{too detailed/handwavy?}

Interactions between attributes are expressed as Horn clauses. An example
costraint might look like this: ``if the manufacturer is Apple, then the
CPU must be either a PowerPC G3 or a PowerPC G4'', which can be encoded
into MILP form as:
%
$$ (1 - x_\text{Apple}) + x_\text{G3} + x_\text{G4} \ge 1 $$
%
Note that mutual exclusivity between all binary variables corresponding to an
attribute ensures that only one of $x_\text{G3}$ and $x_\text{G4}$ can be 1.
The dataset includes constraints between the following attributes: manufacturer
$\to$ type, manufacturer $\to$ CPU, type $\to$ RAM amount, type $\to$ storage
size, type $\to$ monitor size. In total we encoded 16 Horn constraints, each
with one head and multiple options in the body. We do not report the full
list due to space limitations.

\begin{table}
    \centering
    \begin{tabular}{ccc}
        {\bf Attribute} & {\bf Type} & {\bf Values} \\
        \hline \hline
        Type & discrete & 3 \\
        Manufacturer & discrete & 8 \\
        CPU model & discrete & 37 \\
        RAM amount & discrete & 10 \\
        HD size & discrete & 10 \\
        Monitor size & discrete & 8 \\
        Price & continuous & --
    \end{tabular}
    \caption{\label{tab:pcdataset} Statistics for the PC dataset.}
\end{table}
